---
title: "Project 3: Joining CSV files"
output:
  html_document:
    pandoc_args: [
      "+RTS", "-K64m",
      "-RTS"
    ]
---
For analysis, we are using two databases. One has US census data from 2010 by zipcode. The other has US income tax information from 2010 by zip code. In analysis, we will employ joins and piping techniques to wrangle the data in order to produce interesting visualizations.

###Reproducing Our Work:
1. Create a GitHub repository for the project.
2. Create the folders for your project.
3. Find a csv file for statistical comparisons. Our CSV files were found on www.data.gov.
4. Import CSV into SQL Developer
5. Build an html file to document these steps.

Before generating plots, we run the 'require.r' file to make sure the necessary packages are in place.
```{r}
source("../require.R")
```

###Summary and Subset:
The SQL used on the table selects all census and income information. A summary and head are displayed for each data frame.
```{r, echo=TRUE}
source("../01 Data/summary.R", echo = TRUE)
```

###Plot #1:
First, the income and census data frames were joined by zip code using a dpylr inner join function. Then, two new columns were added to the new table. We found the percentage of single mother households in each zip code, and then made another new column to analyze them by percentile. (singleMotherP and mother_percentile)

Next, we found the average salaries and wages reported on each income tax return by the population size in each zip code. (avgIncome)

We then classified each zip code by its relative population size using the cumulative distribution of the size of the population. The categories "Small", "Medium", and "Large" represent the smallest third, medium third, and largest zip codes by population size percentile.

Finally, we mutated the average income column into percentiles for analysis and facet wrapped them for easy viewing.

The first interesting thing about this plot is that there is a solid correlation between zip code population size and average salaries/wages per individual in that population. Secondly, we notice that the large zip code areas appear to have more single mother households than the small zip code areas. The medium size zip codes are perfectly in the middle. 

Finally, it appears that zip codes with more single mothers tend to have relatively smaller wages/salaries per individual - note the heavy grouping of points in the top left of each zip code size grouping. Also note here is that there is a shift toward higher relative salary AND higher relative single mother house holds per zip code the larger the zip code is. 

Leading Questions About this Data:
1. Are single mother households forced to take lesser paying jobs to take care of their children? (The classic career for child sacrifice)

2. Why does increasing zip code population size correlate with increased single mother households AND increased salaries per person?

```{r}
source("../02 Data Wrangling/plot1.R", echo = TRUE)
```